Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Edited cluster analysis section

We then explored "features" in our data. There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques. We used hierarchical cluster analysis on our quality and function variables. We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

@@

Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$. Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward agglomeration algorithm.

@@

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters. Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*). The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygynous*). Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).

The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as an organizational functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).

>>>>

We then explored "features" in our data. There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques. We used hierarchical cluster analysis on our quality and function variables. We computed distance between our binary vectors using both $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

@@

@@

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to material, social, and reproductive success, and one that comprised mostly variables indicating the relationship between leaders and their groups, which we term *social orientation*. Within the latter top-level cluster we identify two interpretable though weakly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a subcluster mainly comprising variables related to the ability to provide social benefits, and a subcluster mainly comprising variables related to the ability to impose social costs.

The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*). Within the management cluster, we identified a mediation subcluster (e.g., *Resolve conflict*, *Provide counsel*), and a an organization subcluster (e.g., *Organize cooperation*, *Military command*).

In naming the clusters and subclusters, we gave extra weight to variables that were strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes we identified.
